Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-22 fmics/paper.tex
Start drafting first paragraphs of Section 4

\section{Verification of SRS Implementation using HOL}
\label{sec-verification}

Calculation of the running sums is a critical component of BLM. In case of no dangerous loss, BLM should not inhabit the beam. If it fails, it generates a false alarm and three hours will be lost to recover the previous situation. Such false alarms compromise the availability of the system. Moreover, in case of a dangerous loss, BLM it has to inhibit the beam permit. If it fails there will be at least 30 days of downtime. 
Any error in this computation would either compromise the availability of the system (unnecessary request for beam dump) or safety of the system (nor requesting a necessary beam dump). The current practice to check the correctness of the SRS is by simulation and testing in which (explanation of how such a test gets performed).

In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.

The formal verification effort described in this paper is based on
 the use of mechanised logical deduction, or theorem proving, to
 demonstrate that desired properties of a system are logically implied
 by a formal model of the system.  Our approach to theorem-proving
 uses the HOL4 open source software tool, which was developed
 initially at the University of Cambridge, but now by an international
 team~\cite{HOL4}.  HOL4 partially automates the task of proving theorems of
 higher-order logic [ref], a formal logic with a similar expressive
 power to set theory that is widely used for formalising hardware and
 software models and statements about them.  Typically, a very large
 portion of the logical deduction is performed automatically by the
 HOL4 system leaving the user to provide the high level proof
 strategy.  The implementation of the HOL4 system is based on Milner's
 LCF approach [ref] and consists of a small “kernel” that provides a
 very high level of confidence in the logical soundness of the
 verification results obtained using the system. The work described
 here could have been done using other systems such as HOL Light
 [ref], Isabelle/HOL~\cite{Isabelle}, ProofPower [ref], PVS [ref] or Coq
 [ref]. The first three of these support essentially the same version
 of higher-order logic as HOL4, whilst PVS and Coq support more
 powerful logics. <What about ACL2?>

 While offering less “push button” automation than many other kinds of
 formal verification such as model-checking, machine-assisted theorem
 proving is the most appropriate approach to verifying the BLM SRS,
 since it is unlikely that the correctness theorems, which need lemmas
 proved by manually guided mathematical induction, could be generated
 automatically.

>>>>

\section{Verification of the SRS Implementation using HOL4}
\label{sec-verification}

%The calculation of running sums is a critical component of the BLMS.
%If there is no dangerous loss, the BLMS should not inhibit the beam.
%A false alarm costs at least three hours to resume operation, and compromises the availability of the system.
%However, if there is a dangerous loss, the BLMS must inhibit the beam.
%A quench due to dangerous loss costs at least 30 days of downtime.
%Thus, any error in the calculation compromises either the availability or the safety of the system.
%The current practice to check the correctness of the SRS is by simulation and testing.
%(isn't all this in the introduction now? is this the right place to repeat it/give more details?)
%In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.

Our formal verification effort uses mechanised logical deduction, or \emph{theorem proving}.
In general, theorem proving is used to show that desired properties of a system are logically implied by a formal model of the system.
We use the HOL4 open source software tool~\cite{HOL4,DBLP:conf/tphol/SlindN08}, which was developed initially at the University of Cambridge, but now by an international team.
HOL4 enables the construction of theories in higher-order logic~\cite{DBLP:journals/jsyml/Church40}, a formal logic with a similar expressive power to set theory that is widely used for formalising hardware and software models and statements about them.
The implementation of HOL4 uses Milner's LCF approach~\cite{Milner:1972:LCF:891954}: a small ``kernel'' implementing the primitive rules of the logic, and convenient derived rules and tactics implemented in terms of the kernel.
Every theorem ultimately comes from the kernel, and this fact provides high assurance of the logical soundness of the verification results obtained using the system.

HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
The work described here could have been done using other systems such as HOL Light [ref], Isabelle/HOL~\cite{Isabelle}, ProofPower [ref], PVS [ref] or Coq [ref].
The first three of these support essentially the same version of higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
<What about ACL2?>
While offering less ``push-button'' automation than many other kinds of formal verification such as model-checking, machine-assisted theorem proving is the most appropriate approach to verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
